Easy as Pi: A Network Coding Raspberry Pi Testbed

### 2016-06-06 measurements.tex

In \ref{fig:enc_good_rasp1_gen_gf2}, \ref{fig:enc_good_rasp1_gen_gf256},
\ref{fig:enc_good_rasp2_gen_gf2} and \ref{fig:enc_good_rasp2_gen_gf256};
it can be seen the goodput trends of 5 codes as a function of
the generation size: \ac{RLNC}, \ac{SRLNC} with $d = [0.02,\ 0.1]$
and \ac{TSNC} with two parameters that we mention. The first of them,
$steps$, is the number of density changes in the system, if any.
The number of changes is equal to the number of regions minus 1, making
$steps = \#regions - 1$ and $dep = e$. With these definitions, we first
evaluate \ac{TSNC} with no changes, e.g. $steps=0$ and the two amounts
for dependent packets mentioned earlier. In this way the budget is
either $g + 8$ or $g + 16$. This means that we tolerate at most 8 or 16
packets regardless of the generation. The goal with setting first
\ac{TSNC} with these values is to observe first the sole effect of the
budget in the code since $steps = 0$ implies no density changes during
the whole process.

For the plots in the mentioned figures from the encoder goodput measurements,
\ac{RLNC} presents the worst performance in terms of goodput and
\ac{TSNC} with $dep = 16$, the best regardless of the \ac{Raspi} model.
Given that the processing time depends in the amount of coded packets
required to create, \ac{RLNC} is the slowest to process since it must
use all of the $g$ original packets. Later, sparse codes process the data
at a larger rate since less packets are being mixed when creating a
coded packet. The caveat of these schemes is that the sparser the code,
the more probable the ocurrences of \ac{l.d.} packets is. So, basically
the sparser the codes, then more overhead due to transmissions of \ac{l.d.}
packets they have since it is more likely to generate the same set of
packets. Excluding the coding coefficients overhead, the overhead due to
transmissions of \ac{l.d.} packets can be very high.
For example, if we consider \ac{TSNC} with $dep=16$ and $g = 16$, the budget
in this case permits to send up to 32 packets which is $2\times$ the generation
size for an overhead of $100\%$ excluding the one from the coding
coefficients. This happens because \ac{TSNC} is allowed to add too much
redundancy in this case, which permits it to use a, single in this case,
very low code density $d$. For \ac{RLNC} this is not the case, since the
occurrence of \ac{l.d.} coded packets is low. Even for $GF(2)$, the
average amount of redundancy has proven to be 1.6 packets after $g$ have
been transmitted, much less than the cases where sparse codes are
utilized. Overall, we observe that there is a trade-off between goodput
and \ac{l.d.} coded packets transmission overhead.

For Figs.~\ref{fig:enc_good_rasp1_packet_gf2},
\ref{fig:enc_good_rasp1_packet_gf256}, \ref{fig:enc_good_rasp2_packet_gf2}
and \ref{fig:enc_good_rasp2_packet_gf256}; we see the packet
size effect in the encoding goodput in both \ac{Raspi} models and for
a fixed generation size. For \ac{TSNC} in this case, we allow for 5 density
changes during the generation transmission and again consider the same
budgets as before. In all the cases, we clearly observe that there is an
optimal packet size to operate with. When reviewing the specifications
of the \ac{Raspi} 1, it uses a L1 cache of 16 KiB. Hence, the trends
in the packet size can be explained as follows: For low packet sizes,
the data processing does not overload the \ac{CPU} of the \ac{Raspi}, so
the goodput progressively increases given that we simply process more
data as the packet size increases. However, after a certain packet size,
the goodput gets affected since the cache starts to saturate. The packets
towards the cache needs to have more processing. Also important, the
\ac{CPU} needs to allocate computational resources to other processes from
the different tasks running in the \ac{Raspi}. Beyond this critical packet
size, the \ac{CPU} just queues processing which incurs in larger delay
reducing the goodput.

If we consider the previous effect when reviewing the trends in the
mentioned figures, in all the models we observe that the maximum does not
come at 16 KiB but at a smaller value depending on the model, field size
and coding scheme considered. The reason is that given there are various
tasks from other applications for proper functioning running in the
\ac{Raspi} at the same time as coding, the cache space is filled also
with the data from these tasks, thus diminishing the cache space available
for coding.

For the \ac{Raspi} 1 model, we observe in
Fig.~\ref{fig:enc_good_rasp1_packet_gf2} that the critical
packet size for \ac{RLNC} using $GF(2)$ occurs at 1 KiB whereas for the
sparse codes is close to 4 KiB in the most of the cases for $GF(2)$. This
difference takes place since the sparse codes are much less complex to
process than \ac{RLNC}, which turns into less data loading in the
cache for doing computations. However, for \ac{RLNC} given that it is
a dense code since $d \to 1$, \ac{RLNC} packets load much more data to
the cache quite easily making this effect visible for smaller packet sizes.
Fig.~\ref{fig:enc_good_rasp1_packet_gf256}, although the ideal packet
size remains the same for \ac{RLNC}, for sparse codes it locates closer to
8 KiB.
%
\begin{figure*}
    \centering
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_generation_size_Rasp_v2_encoder_Binary_1600.eps}
        \caption[]%
        {{\small Goodput vs. Generation size for $q = 2$}}
        \label{fig:enc_good_rasp2_gen_gf2}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_generation_size_Rasp_v2_encoder_Binary8_1600.eps}
        \caption[]%
        {{\small Goodput vs. Generation size for $q = 2^8$}}
        \label{fig:enc_good_rasp2_gen_gf256}
    \end{subfigure}
    \vskip\baselineskip
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_symbol_size_Rasp_v2_encoder_Binary_16.eps}
        \caption[]%
        {{\small Goodput vs. Packet size for $q = 2,\ g=16$}}
        \label{fig:enc_good_rasp2_packet_gf2}
    \end{subfigure}
    \quad
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_symbol_size_Rasp_v2_encoder_Binary8_16.eps}
        \caption[]%
        {{\small Goodput vs. Packet size for $q = 2^8,\ g=16$}}
        \label{fig:enc_good_rasp2_packet_gf256}
    \end{subfigure}
    \caption[]
    {\small Encoder goodput measurements for the \ac{Raspi} 2}
    \label{fig:enc_good_rasp2}
\end{figure*}
%
\subsubsection{Decoding}

>>>>

In \ref{fig:enc_good_rasp1_gen_gf2}, \ref{fig:enc_good_rasp1_gen_gf256},
\ref{fig:enc_good_rasp2_gen_gf2} and \ref{fig:enc_good_rasp2_gen_gf256};
it can be seen the goodput trends of 5 codes as a function of
the generation size: \ac{RLNC}, \ac{SRLNC} with $d = [0.02,\ 0.1]$
and \ac{TSNC} with an extra parameter that we mention. We define
$C$ as the number of density changes in the \ac{TSNC} transmission
process, if any. The number of changes is $C = \#regions - 1$. With
this definitions, we first evaluate \ac{TSNC} with no changes, e.g.
$C = 0$ and the two amounts for dependent packets mentioned earlier.
This means that we tolerate at most 8 or 16 packets regardless of the
generation. The goal with setting first \ac{TSNC} with these values is to
observe first the sole effect of the budget in the code since $C = 0$
implies no density changes during the whole process.

For the plots in the mentioned figures from the encoder goodput measurements,
\ac{RLNC} presents the lowest performance in terms of goodput and
\ac{TSNC} with $e = 16$, the highest regardless of the \ac{Raspi} model.
Given that the processing time depends in the amount of coded packets
required to create, \ac{RLNC} is the slowest to process since it must
use all of the $g$ original packets. Later, sparse codes process the data
at a larger rate since less packets are being mixed when creating a
coded packet. The caveat of these schemes is that the sparser the code,
the more probable the ocurrences of \ac{l.d.} packets is. So, basically
the sparser the codes, then more overhead due to transmissions of \ac{l.d.}
packets since it is more frequent to generate similar set of packets.
Excluding the coding coefficients overhead, the overhead due to
transmissions of \ac{l.d.} packets might be high for the sparse schemes.
For example, if we consider \ac{TSNC} with $e = 16$ and $g = 16$
in Fig.~\ref{fig:enc_good_rasp1_gen_gf2}, the budget in this case permits
to send up to 32 packets which is $2\times$ the generation
size for an overhead of $100\%$ excluding the overhead from appending
the coding coefficients. This happens because \ac{TSNC}
has been allowed to add too much redundancy in this case. Given that $C=0$,
this permits it to use a single very low code density $d$. For \ac{RLNC},
this is not the case since the occurrence of \ac{l.d.} coded packets is
small, because all coding coefficients are used. Even for $GF(2)$, the
average amount of redundant packets for \ac{RLNC} has proven to be 1.6
packets after $g$ have been transmitted \textbf{[Include a reference]},
but less than the cases where sparse codes are utilized. Overall, we
observe that there is a trade-off between goodput and \ac{l.d.} coded
packets transmission overhead.

For Figs.~\ref{fig:enc_good_rasp1_packet_gf2},
\ref{fig:enc_good_rasp1_packet_gf256}, \ref{fig:enc_good_rasp2_packet_gf2}
and \ref{fig:enc_good_rasp2_packet_gf256}; we see the packet
size effect in the encoding goodput in both \ac{Raspi} models and for
a fixed generation size. For \ac{TSNC} in this case, we allow for 5 density
changes during the generation transmission and again consider the same
budgets as before. In all the cases, we observe that there is an
optimal packet size to operate with. When reviewing the specifications
of the \ac{Raspi} 1, it uses a L1 cache of 16 KiB. Hence, the trends
in the packet size can be explained in the following way: For low packet
sizes, the data processing does not overload the \ac{CPU} of the
\ac{Raspi}, so the goodput progressively increases given that we simply
process more data as the packet size increases. However, after a certain
packet size, the goodput gets affected since the cache starts to saturate.
The packets towards the cache needs to have more processing. Beyond this
critical packet size, the \ac{CPU} just queues processing which incurs in
larger delay reducing the goodput.

If we consider the previous effect when reviewing the trends in the
mentioned figures, in all the models we observe that the maximum is not
at 16 KiB but at a smaller value depending on the model, field size
and coding scheme considered. The reason is that the \ac{CPU} needs to
allocate computational resources to other processes from the different tasks
running in the \ac{Raspi}. Then, given there are various
tasks from other applications for proper functioning running in the
\ac{Raspi} at the same time as coding, the cache space is filled also
with the data from these tasks, thus diminishing the cache space available
for coding operations.

For the \ac{Raspi} 1 model, we observe in
Fig.~\ref{fig:enc_good_rasp1_packet_gf2} that the critical
packet size for \ac{RLNC} using $GF(2)$ occurs at 1 KiB whereas for the
sparse codes is close to 8 KiB in the most of the cases for $GF(2)$. This
difference takes place since the sparse codes mix less packets their
process than \ac{RLNC}, which turns into less data loading in the
cache for doing computations. For a density of $d = 0.1$, packet size of
$B = 8$ KiB and $g = 16$ packets, we observe that roughly $\ceil{gd} = 2$
packets are inserted in the cache when calculating a coded packet with this
sparse code. Loading this into the cache, stands for 16 KiB which is the
cache size. A similar effect occurs for the other sparse codes. However, for
\ac{RLNC} given that it is a dense code since $d \to 1$, \ac{RLNC}
packets load data from all the $g = 16$ coding coefficients, which accounts
for the 16 KiB of the cache size. In
Fig.~\ref{fig:enc_good_rasp1_packet_gf256}, although the ideal packet
size remains the same for \ac{RLNC} and the sparse codes, the final goodput
is lower due to the field size effect.
%
\begin{figure*}
    \centering
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_generation_size_Rasp_v2_encoder_Binary_1600.eps}
        \caption[]%
        {{\small Goodput vs. Generation size for $q = 2$}}
        \label{fig:enc_good_rasp2_gen_gf2}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_generation_size_Rasp_v2_encoder_Binary8_1600.eps}
        \caption[]%
        {{\small Goodput vs. Generation size for $q = 2^8$}}
        \label{fig:enc_good_rasp2_gen_gf256}
    \end{subfigure}
    \vskip\baselineskip
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_symbol_size_Rasp_v2_encoder_Binary_16.eps}
        \caption[]%
        {{\small Goodput vs. Packet size for $q = 2,\ g=16$}}
        \label{fig:enc_good_rasp2_packet_gf2}
    \end{subfigure}
    \quad
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_symbol_size_Rasp_v2_encoder_Binary8_16.eps}
        \caption[]%
        {{\small Goodput vs. Packet size for $q = 2^8,\ g=16$}}
        \label{fig:enc_good_rasp2_packet_gf256}
    \end{subfigure}
    \caption[]
    {\small Encoder goodput measurements for the \ac{Raspi} 2}
    \label{fig:enc_good_rasp2}
\end{figure*}
%
The effects mentioned for the \ac{Raspi} 1 were also observed for
the \ac{Raspi} 2.
\subsubsection{Decoding}
